---
title: "Chocolate Bar Rating"
output: 
  ioslides_presentation:
  widescreen: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# initialize libraries 
library(tidyverse)
library(DT)
library(GGally)
library(ggridges)
# Read the data 
chocolate <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')
candy_data <- read_csv("../data/candy_data.csv")

# Change format of some columns
chocolate$cocoa_percent <- gsub("%","",as.character(chocolate$cocoa_percent))
chocolate$cocoa_percent <- as.numeric(chocolate$cocoa_percent)
```

## Table of Content: 
- Introduce the dataset and scenario.

- Exploratory Data Analysis.

- External Data Set Analysis.  

- Key investigations results from the data.

## Introduction 

Chocolate is the most popular candy in the world. The world consumption is estimated to be at least 7.2 mill metric tons/year, so chocolate has a strong business market. However, not all chocolate bars are created equal.

This dataset contains expert ratings of over 1,700 individual chocolate bars, along with other information.

## Data Summary
This rating were compiled by Brady Belinsky, founding member of the Manhattan chocolate society

- Data collection duration  2006-2021

- From 580 company

- 67 different country

The dataset contains 2530 entry each one has 10 attribute list as following:

- Numeric: Ref, Review Year, Cocoa Percent, Rating.

- Character: Company Name, Company Location, Bean Origin, Bar Name, Ingredients, Characteristic

## Data Table
```{r echo=FALSE}
chocolate %>% 
  select(company_location, country_of_bean_origin, cocoa_percent, rating) %>%
  datatable(options = list(pageLength = 7))
```


## Flavors of Cacao Rating System


- 5= Elite (Transcending beyond the ordinary limits)

- 4= Premium (Superior flavor development, character and style)

- 3= Satisfactory(3.0) to praiseworthy(3.75) (well made with special qualities)

- 2= Disappointing (Passable but contains at least one significant flaw)

- 1= Unpleasant (mostly unpalatable)

## Data Quality

- Missing Data :87 missing value (all under ingredients)

- Inconsistency : some values inconsistent (combined country and bar names)

- Cocoa Percent is a character, not a numeric value, percent sign removed

- Check the names and remove the spaces and line breaks 

## Questions Navigating Our Analysis

- What country/countries are manufacturing the highest-rating chocolate bars?

- What country/countries are growing the highest-rating beans? 

- What is the trend between the rating and cocoa percentage?

- What ingredients are found most among the highest-rating chocolate bars?

- Does the time in which the review was made affect the rating? 

## Rating Density Plot
```{r, echo=FALSE}
ggplot(chocolate, aes(rating)) + 
  geom_density()
```

## Highest Rating Company Location 
```{r, echo=FALSE}
names(sort(table(chocolate$company_location), decreasing = TRUE))[1:10] -> top_countries
top_rating <- seq(3,4,0.25)
chocolate %>%  
  filter(company_location %in% top_countries,
         rating %in% top_rating) %>% 
  group_by(rating) %>% 
  ggplot(aes(company_location, fill = company_location)) +
  geom_bar() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x = element_blank()) +
  facet_wrap(vars(rating)) 
```

## Highest Rating Bean Origin Location
```{r, echo=FALSE}
names(sort(table(chocolate$country_of_bean_origin), decreasing = TRUE))[1:10] -> top_bean_countries
chocolate %>%  
  filter(country_of_bean_origin %in% top_bean_countries,
         rating %in% top_rating) %>% 
  group_by(rating) %>% 
  ggplot(aes(country_of_bean_origin, fill = country_of_bean_origin)) +
  geom_bar() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x = element_blank()) +
  facet_wrap(vars(rating)) 
```

## Cocoa% Rating Densities 
```{r, echo=FALSE}
chocolate %>% 
  group_by(rating) %>% 
  ggplot(aes(y = as.character(rating),  x = cocoa_percent, fill = stat(quantile))) + 
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient") +
  scale_fill_brewer(name = "") 
```

## Hypothesis Test
Population: 4 rated chocolates. 

H_0: 70.25 - mew = 0

- Sample (n): 112

- Mean (x_bar): 70.70

- Standard deviation (s): 3.37

X_bar ~ N(70.70, (3.37^2)/112)

t_stat = (x_bar - mew_0)/SEM = (-1.41, 1.41)

p = 0.159 -> Cannot Reject Null Hypothesis


## Most Used Ingredients Rating
```{r, echo=FALSE}
names(sort(table(chocolate$ingredients), decreasing = TRUE))[1:4] -> top_ingredients

# plot top ingredients and their cocoa percent and rating graphs:
chocolate %>% 
  filter(ingredients %in% top_ingredients) %>% 
  ggplot(aes(x = as.character(rating))) +
  geom_bar() +
  facet_wrap(vars(ingredients))
```


## Rating Density Based on Date 
```{r, echo=FALSE}
chocolate %>% 
  group_by(rating) %>% 
  ggplot(aes(y = as.character(rating),  x = review_date, fill = stat(quantile))) + 
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient") +
  scale_fill_brewer()
```

## External Data Set 
Main goal: Investigate whether the high ranked candies contain chocolate or not.

The dataset contains 85 entries and 13 variables listed as follows:

- Logical: chocolate, fruity, caramel, peanutalmondy, nougat, crispedricewafer, hard, bar, pluribus
- Continuous: sugarpercent, pricepercent, winpercent
- Character: competitorname

## Table 
```{r, echo=FALSE}
candy_data %>%
  arrange(-winpercent) %>% 
  select(competitorname, winpercent, chocolate) %>% 
  datatable(options = list(pageLength = 7))
```

## Take Aways

- Rating Density: Negative skew, median = 3.25.

- Highest Rating Company Location: USA for the win? 

- Highest Rating Bean Origin Location: Venezuela and Peru lead the race with Dominican Republic slightly behind. 

- Cocoa Percentage Density: Top rating bar have cocoa percent between 60 and 80. 

- Most Used Ingredients Rating: Beans, sugar, and cocoa butter make up most bars. 

- Rating Density Based on Date: Less/No low rating bars produced recently. 
